\section{Security Considerations}
\subsection{Threat Model}
This format is designed to protect files at rest and in transport from accidental disclosure.
%
Using authenticated encryption in individual segments mirrors solutions like Transport Layer Security (TLS) as described in~\cite{RFC5246} and prevents undetected modifications of segments.

\subsection{Selection of Key and Nonce}
The security of the format depends on attackers not being able to guess the encryption key (and to a lesser extent the nonces).
%
The encryption key and nonce SHOULD be generated using a cryptographically-secure pseudo-random number generator and MUST NOT be repeated with the same encryption key.
%
This makes the chance of guessing a key vanishingly small.
%
Additional security can be provided by using `Associated Data' when encrypting a file.
%
This data must be used to decrypt the data, although it is not part of the encrypted file \cite{RFC8439}

\subsection{Message Forgery}
The encrypter's public key component is required to calculate the encryption key used for the header in the Curve25519 function. 
If this public key is not specified in the header directly, it must be obtained from the encrypter, and can then be used to validate the origin of the file.
The encrypted header part is authenticated using the Poly1305 algorithm over the header data.
%
Using ChaCha20-ietf-Poly1305, the Poly1305 algorithm produces an authentication tag for each encrypted segment of the encryoted data.

\subsection{No File Updates Permitted}
Implementations MUST NOT update encrypted files.
%
Once written, a section of the file must never be altered.
